### 2021-01-14 docs/cropland_labels.Rmd
Finish labels

---
title: "Cropland labels"
author: "Lei Song"
date: "1/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Object-based cropland labels
```{r ecozones}
ecozones <- st_read('data/ecozones/ecozones_shp.shp', 
                    quiet = T) %>% 
  group_by(PHY_REG) %>% summarise() %>% 
  setNames(c('zone', 'geometry'))
st_write(ecozones, 'data/geoms/ecozones.geojson')
```

### Validation
Choose out 100 grids for evaluation. These 100 grids will be used to evaluate the labeller as well.

```{r val_grids}
set.seed(1000)
vals_grid <- ecozones %>% 
  st_sample(size = rep(10, nrow(.))) %>% 
  st_sf() %>% mutate(id = 1:nrow(.))
```

### Calibration
First, skip the buffered area around validation grids.

```{r cal_grids}
vals_grid_buffer <- vals_grid %>% 
  st_buffer(dist = 0.5, endCapStyle = 'SQUARE') %>% 
  st_union()
ecozones_dif <- st_difference(ecozones, vals_grid_buffer)
```