### 2021-01-14 docs/run_dl_aws.Rmd

---
title: "Run deep learning on AWS instance"
author: "Lei Song"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    highlight: pygments
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Launch instance

## Parameters setting

## Result visualization