### 2021-01-16 docs/run_dl_aws.Rmd

---
title: "Run deep learning on AWS instance"
author: "Lei Song"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    highlight: pygments
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Launch instance
The information of instance that we used:

- AMI ID: `ami-0643df3b347740189`, which is our customized deep learning AWS instance.
- Instance type: `g3.8xlarge`.
- Security group: `labeller-security`.

Before start, install `jq`.
Then use the scripts in `pytorch_planet/aws_tools/create_spot_instance.sh` using the template:
```
cd /path/to/pytorch_planet/folder
chmod 777 ./aws_tools/create_spot_instance.sh
./aws_tools/create_spot_instance.sh <ami_id> <instance_type> <security_group_id> <new_instance_name> <spot_type> <valid_until> <key_name>
```
For example:
```
cd /path/to/pytorch_planet/folder
chmod 777 ./aws_tools/create_spot_instance.sh
./aws_tools/create_spot_instance.sh ami-0643df3b347740189 g3.8xlarge sg-0a8bbc91697d6a76b lc_tz persistent 2021-01-20T23:00:00 lsong-keypair
```

## Parameters setting
After successfully launched the instance, then it is ready to get into the instance to work.

The most essential work are arguably to clone project and set jupyter notebook:

- STEP 1: Set up GitHub token (because the repo is private). More details could be found in this [tutorial](https://docs.github.com/en/free-pro-team@latest/github/authenticating-to-github/generating-a-new-ssh-key-and-adding-it-to-the-ssh-agent).

  ```
  ssh-keygen -t ed25519 -C "your_email@example.com"
  eval "$(ssh-agent -s)"
  ssh-add ~/.ssh/id_ed25519
  vim ~/.ssh/id_ed25519.pub
  ```
  Copy out the ssh key, and type `:q` to exit, and then go to GitHub page to add into the records.

- STEP 2: Clone the repo into instance:

  ```
  git clone git@github.com:agroimpacts/pytorch_planet.git
  ```

  If necessary, use `git checkout` to switch branch and use `git pull` to grab the updates.

- STEP 3: Set up jupyter notebook

  ```
  jupyter notebook password
  ```
  to set a log in password for jupyter notebook, then
  ```
  jupyter notebook
  ```
  replace the private IP with the public IP to open it in browser. If you can log in without issues, then it is ready to go.

## Result visualization