### 2022-01-24 docs/ecrs_2022_poster_imgs.Rmd
Add poster for ECRS 2022

---
title: "ecrs_2022_poster"
author: "Lei Song"
date: "1/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(terra)
library(stars)
library(tmap)
library(ggpubr)
library(gridExtra)

pred_path <- here(file.path('results/dl/prediction',
                           'unet_k15_only_clr_200epc',
                           'landcover_north_withwl.tif'))
color_tb <- data.frame(
    id = 1:8,
    type = c("Cropland", "Forest", 
             "Grassland", "Shrubland", 
             "Water", 'Built-up', "Bareland", 
             'Wetland'),
    color = c('#ff7f00', '#074702', '#b2df8a',
              '#33a02c', '#3347f9', '#5a6057',
              '#fdbf6f', '#0ea591'))
pred <- read_stars(pred_path)

# plot
tmap_mode("plot")
pred_final <- tm_shape(pred) +
    tm_raster("landcover_north_withwl.tif", 
              title = "Land cover types",
              n = 8, style = 'cat',
              labels = color_tb$type,
              palette = color_tb$color) +
    tm_layout(frame = F,
              legend.show = F,
              outer.margins = c(0.02, 0.01, 0, 0.01),
              inner.margins = 0,
              bg.color = 'transparent')
pred_final <- tmap_grob(pred_final)

# Get legend
r_df <- data.frame(x = 1:8, 
                   y = 1:8,
                   rst = as.factor(1:8))
legend_lc <- ggplot() + 
    geom_raster(
        data = r_df, 
        aes(x = x, y = y, 
            fill = rst)) +
    scale_fill_manual('Land cover',
                      values = color_tb$color, 
                      breaks = 1:8,
                      labels = color_tb$type) +
    theme_minimal() +
    theme(legend.position = 'right',
          legend.text = element_text(color = 'black'),
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          plot.margin = unit(rep(0, 4),"cm"),
          plot.background = element_rect(fill = "transparent"))+
    guides(fill = guide_legend(ncol = 1))
legend_lc <- get_legend(legend_lc)
pred_final_plt <- ggarrange(pred_final, legend_lc, NA,
                        ncol = 3, widths = c(5, 1, 2))
ggsave(here('docs/figures/ecrs_results.png'),
       pred_final_plt, 
       width = 8, height = 6, dpi = 600)
```

```{r}
library(raster)
library(glue)
library(here)
library(sf)
library(nngeo)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
library(sf)
library(rasterVis)
library(ggnewscale)
library(cowplot)
library(ggpubr)

# Read boundaries
study_area <- read_sf(here('data/geoms/study_area.geojson')) %>% 
    st_remove_holes()
nicfi_tiles <- read_sf(here('data/geoms/tiles_nicfi_north.geojson')) %>% 
    mutate(`NICFI tile` = 1)

# Make the main figure
main_fg <- ggplot(nicfi_tiles) +
    geom_sf(fill = NA, 
            aes(color = "grey"), 
            lwd = 0.5) +
    geom_sf(data = study_area, fill = NA, 
            aes(color = 'red', lty = 'solid')) +
  scale_color_identity(name = '',
                         breaks = c('black', 'red'),
                         labels = c('NICFI tiles', 'Study area'),
                         guide = 'legend') +
    scale_linetype_identity(name = '',
                            breaks = c('dotted', 'solid'),
                            labels = c('NICFI tiles', 'Study area'),
                            guide = 'legend') +
    theme_minimal() + 
    theme(legend.position = 'none',
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.margin = unit(rep(0, 4),"cm"))
ggsave(here('docs/figures/ecrs_tiling.png'),
       main_fg, 
       width = 6, height = 6, dpi = 600)

id_selected <- '1226-999'
# Read tiles
nicfi_tiles <- read_sf(
    here('data/geoms/tiles_nicfi_north.geojson'))

# Sub-tiling
indices <- matrix(1:64, 8, 8)
indices <- indices[, ncol(indices):1]
sub_tiling <- nicfi_tiles %>% 
    filter(tile == id_selected) %>% 
    st_make_grid(n = c(8, 8)) %>% 
    st_sf() %>% 
    mutate(index = as.vector(indices))
trains_inside <- sub_tiling %>% 
    filter(index %in% (read.csv(
    here('results/north/dl_catalog_train.csv'),
    stringsAsFactors = F) %>% 
    filter(tile == id_selected) %>% 
    pull(index))) %>% 
    mutate(Usage = 'Train')
valids_inside <- sub_tiling %>% 
    filter(index %in% (read.csv(
    here('results/north/dl_catalog_valid.csv'),
    stringsAsFactors = F) %>% 
    filter(tile == id_selected) %>% 
    pull(index))) %>% 
    mutate(Usage = 'Validate')
# Labels of each sub-tile
sub_tiles <- rbind(trains_inside, valids_inside) %>% 
    arrange(index); rm(trains_inside, valids_inside)

# Sub-tiling figure
sub_tiling_fg <- ggplot(sub_tiling) + 
    geom_sf(data = sub_tiles, aes(fill = Usage),
            color = 'black', lwd = 0.3) +
    scale_fill_manual(values = c('forestgreen', 'orangered')) +
    geom_sf(color = 'black', fill = NA, lwd = 0.3) + 
    geom_sf_text(aes(label = index), size = 10) +
    theme_minimal() +
    theme(legend.position = 'bottom',
          legend.title = element_blank(),
          legend.text = element_text(size = 30),
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          plot.title = element_text(hjust = -0.15, vjust = -6),
          plot.margin = unit(rep(0, 4),"cm"))

ggsave(here('docs/figures/ecrs_subtiling.png'),
       sub_tiling_fg, 
       width = 6, height = 6, dpi = 600)

sub_tiles <- sub_tiles %>% filter(index == 12)

# Weak labels
labels_esb <- raster(here('data/north/lc_labels_north_mask.tif')) %>% 
        crop(sub_tiles)

labels_guess <- lapply(sub_tiles$index, function(id){
    prefix <- glue('results/north/guess_labels/guess_{id_selected}')
    raster(here(glue('{prefix}_{id}.tif')))
})[[1]]

labels_refine <- lapply(sub_tiles$index, function(id){
    prefix <- glue('results/north/refine_labels/{id_selected}')
    raster(here(glue('{prefix}_{id}_label.tif')))
})[[1]]

labels_dl <- st_crop(pred, st_transform(sub_tiles, 
                                        crs = st_crs(pred)))

# Plot
color_tb <- data.frame(
    id = 1:7,
    type = c("Cropland", "Forest", 
             "Grassland", "Shrubland", 
             "Water", 'Built-up', "Bareland"),
    color = c('#ff7f00', '#074702', '#b2df8a',
              '#33a02c', '#3347f9', '#5a6057',
              '#fdbf6f'))
plot_rst <- function(rst, bry){
    r_df <- as.data.frame(rst, xy = TRUE) %>%
        `colnames<-`(c('x', 'y', 'rst')) %>%
        mutate(rst = as.factor(rst)) %>% 
        na.omit()
    colors <- color_tb %>% filter(id %in% unique(r_df$rst))

    # plot
    ggplot() + 
        geom_raster(
            data = r_df, 
            aes(x = x, y = y, 
                fill = rst)) +
        scale_fill_manual(
            values = colors %>% pull(color), 
            breaks = colors %>% pull(id),
            labels = colors %>% pull(type)) +
        geom_sf(data = bry, 
                color = 'blue', fill = NA,
                lty = 'dotted', lwd = 0.4) +
        theme_minimal() +
        theme(legend.position = 'none',
              axis.title.x = element_blank(),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              axis.title.y = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(),
              plot.margin = unit(rep(0, 4),"cm"))
}

plots_esb <- plot_rst(labels_esb, sub_tiles)

plots_guess <- plot_rst(labels_guess, 
                        sub_tiles %>% 
                            st_transform(crs = crs(labels_guess)))

plots_refine <- plot_rst(labels_refine, 
                        sub_tiles %>% 
                            st_transform(crs = crs(labels_refine)))
plots_dl <- plot_rst(labels_dl, sub_tiles %>% 
                            st_transform(crs = st_crs(labels_dl)))

# Ensemble plots
labels_fg <- ggarrange(plots_esb, plots_guess, plots_refine, 
                       plots_dl, ncol = 4)
ggsave(here('docs/figures/exrs_exp_labels.png'),
       labels_fg, width = 10, height = 4, dpi = 300)
```

```{r}
id_selected <- '1226-1002'
# Read tiles
nicfi_tiles <- read_sf(
    here('data/geoms/tiles_nicfi_north.geojson'))

# Sub-tiling
indices <- matrix(1:64, 8, 8)
indices <- indices[, ncol(indices):1]
sub_tiles <- nicfi_tiles %>% 
    filter(tile == id_selected) %>% 
    st_make_grid(n = c(8, 8)) %>% 
    st_sf() %>% 
    mutate(index = as.vector(indices)) %>% 
    filter(index == 44)
fnames_nicfi <- list.files("/Volumes/elephant/plt_nicfi",
                           pattern = id_selected,
                           full.names = T)
nicfi_s1 <- stack(fnames_nicfi[1]) %>% subset(1:4)
nicfi_s1 <-  nicfi_s1%>% 
    crop(sub_tiles %>% st_transform(crs = crs(nicfi_s1)))

s1_fg <- ggRGB(nicfi_s1, r = 4, g = 3, b = 2, 
               scale = max(values(nicfi_s1)[, 2:4]), 
               stretch = 'hist') +
    theme_minimal() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.margin = unit(rep(0, 4),"cm"))
```